---
title: "ST558 HW4: Hodge Podge"
format: html
editor: visual
---

# Document Details

#### Author: *Smit Miyani*

#### Collaborators: *N/A*

#### Assignment: *HW4*

#### Date: *23JUN24*

#### Purpose

*General practice in working with advanced functions.*\

```{r}
knitr::opts_chunk$set(message = TRUE, warning = TRUE)

#Remove Warnings
```

# Tasks

## Task 1: Conceptual Questions

1.  What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?

    > The purpose of `lapply()`: `lapply()` in R is used to apply a function to each element of a list or vector, returning a list of the results. In the `purrr` package, the equivalent function to `lapply()` is `map()`.

2.  Suppose we have a list called `my_list`. Each element of the list is a numeric data frame (all columns are numeric). We want use `lapply()` to run the code `cor(numeric_matrix, method = "kendall")` on each element of the list. Write code to do this below

    > `lapply(my_list, function(df) cor(df, method = "kendall"))`.

3.  What are two advantages of using `purrr` functions instead of the BaseR `apply` family?

    > 1\. `purrr` functions always return outputs in a predictable format (usually a list), regardless of the input data structure. making it easier to work with results consistently across different scenarios.
    >
    > 2\. `purrr` integrates with other `purrr` functions and the tidyverse ecosystem very well, making it easier to combine operations and make chaining easy thus reduces the need for writing custom helper functions.

4.  What is a side-effect function?

    > A side effect function is a function that not only computes a result but also performs actions that may alter the program's state or interact with the external environment.

5.  Why can you name a variable `sd` in a function and not cause any issues with the `sd` function?

    > Because R keeps track of variables separately within each function. When R sees `sd` inside the function, it looks first at the function's own set of variables before checking if `sd` is a function. This avoids conflicts and lets us use the same name `sd` for a variable without affecting how the `sd` function works globally.

## Task 2: Writing R Functions

### Loading Packages

Reading required tidyverse packages:

```{r}
library(tidyverse)
```

### Question 1: Creating `getRMSE()` function

Creating a function `getRMSE` that takes in vector of observations, and predictions and returns RMSE. Using `…` to have additional arguments for handling of NA values which would be used within the `mean()` function.

```{r}
getRMSE <- function(responses, predictions, ...) {
  # Ensuring the responses and predictions are numeric vectors of the same length
  if (length(responses) != length(predictions) | (!is.numeric(responses)) | (!is.numeric(predictions))) {
    stop("Input vectors 'responses' and 'predictions' must be numeric and of the same length.")
  }
  # Calculating RMSE. 
  rmse <- sqrt(mean((responses - predictions)^2, ...))
  
  return(rmse)
}

```

### Question 2: Testing the `getRMSE()` function

Generating data (code given in assignment) to test the `getRMSE()` function.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

#### Testing with no `NA` values

Testing `getRMSE()` with `resp` and `pred` .

```{r}
getRMSE(responses = resp,predictions = pred)
```

#### Testing with vectors containing `NA_real_`

Replacing two random observations (choosing index at random using `sample()` from 1: `length(resp)` in `resp` with `NA_real_` .

```{r}
set.seed(55)

random_index <- sample(1:length(resp),size = 2)

resp[random_index[1]] <- NA_real_
resp[random_index[2]] <- NA_real_

sum(is.na(resp))
```

##### RMSE without handling `NA`

```{r}
getRMSE(predictions = pred,responses = resp)
```

##### RMSE with an argument for handling `NA`

```{r}
getRMSE(predictions = pred,responses = resp, na.rm = TRUE)
```

### Question 3: Creating `getMAE()` function

Creating `getMAE()` function similar to `getRMSE()`

```{r}
getMAE <- function(responses, predictions, ...) {
  # Ensuring the responses and predictions are numeric vectors of the same length
  if (length(responses) != length(predictions) | (!is.numeric(responses)) | (!is.numeric(predictions))) {
    stop("Input vectors 'responses' and 'predictions' must be numeric and of the same length.")
  }
  # Calculating MAE 
  mae <- mean(abs(responses - predictions), ...)
  
  return(mae)
}
```

### Question 4: Testing the `getMAE()` function

Generating data (code given in assignment) to test the `getMAE()` function.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

```

#### Testing with no `NA` values

Testing `getMAE()` with `resp` and `pred` .

```{r}
getMAE(responses = resp,predictions = pred)
```

#### Testing with vectors containing `NA_real_`

Replacing two random observations (choosing index at random using `sample()` from 1: `length(resp)` in `resp` with `NA_real_` .

```{r}
set.seed(35)

random_index <- sample(1:length(resp),size = 2)

resp[random_index[1]] <- NA_real_
resp[random_index[2]] <- NA_real_

sum(is.na(resp))
```

##### MAE without handling `NA`

```{r}
getMAE(predictions = pred,responses = resp)
```

##### MAE with an argument for handling `NA`

```{r}
getMAE(predictions = pred,responses = resp, na.rm = TRUE)
```

### Question 5: Wrapper Function

Creating `getMetrics()` which calls the helper function

```{r}
getMetrics <- function(responses, predictions,metric = "BOTH", ...) {
  # Ensuring the responses and predictions are numeric vectors of the same length
  if (length(responses) != length(predictions) | (!is.numeric(responses)) | (!is.numeric(predictions)) | (!is.vector(responses)) | (!is.vector(responses))) {
    stop("Input vectors 'responses' and 'predictions' must be numeric and of the same length.")
  }
  # Calculating metrics 
  RMSE<- getRMSE(responses = responses, predictions = predictions,...)
  MAE <- getMAE(responses = responses, predictions = predictions,...)
  
  METRICS <- c(RMSE,MAE)
  names(METRICS) <- c("RMSE","MAE")
  
  if (metric == "BOTH"){
    return(METRICS)
  } else if (metric == "RMSE"){
    return(METRICS[1])
  } else if (metric == "MAE"){
    return(METRICS[2])
  } else {
    "Choose RMSE,MAE or BOTH"
  }
}
```

### Question 6: Testing `getMetrics()`

Generating data (code given in assignment) to test the `getMetrics()` function.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

#### Calling specific metric

##### Calling for RMSE

```{r}
getMetrics(predictions = pred,responses = resp,metric = "RMSE")
```

##### Calling for MAE

```{r}
getMetrics(predictions = pred,responses = resp,metric = "MAE")
```

##### Calling for BOTH

```{r}
getMetrics(predictions = pred,responses = resp,metric = "BOTH")
```

#### Testing with `NA` values

Replacing two random observations similar to previous instance.

```{r}
set.seed(45)

random_index <- sample(1:length(resp),size = 2)

resp[random_index[1]] <- NA_real_
resp[random_index[2]] <- NA_real_

sum(is.na(resp))
```

##### `getMetrics` without handling `NA`

```{r}
getMetrics(predictions = pred,responses = resp)
```

##### `getMetrics` with an argument for handling `NA`

```{r}
getMetrics(predictions = pred,responses = resp, na.rm = TRUE)
```

##### `getMetrics` with incorrect data

Passing in a dataframe instead of a vector.

```{r}
#| error: true
getMetrics(predictions = data.frame(x),responses = resp, na.rm = TRUE)
```

## Task 3: Quering an API and a Tidy-Style Function

Necessary Packages:

```{r}
library(httr)
library(jsonlite)
```

### Question 1: Querying the API and storing as `JSON` .

#### 

Using a URL to query the News API for top headlines using `GET()`, then retrieing raw content, converting it to a character string with `rawToChar()`, and parsing it into a structured R object using `fromJSON()`.

```{r}

url <- "https://newsapi.org/v2/top-headlines?country=gb&q=Euro&from=2024-05-25&to=2024-06-25&category=sports&apiKey=<key_stored_locally>" 

news <- GET(url)
news_data <- rawToChar(news$content)
news_json <- fromJSON(news_data)
```

### Question 2: Returning the dataframe with articles information .

Extracting `news_json$articles` which contains the all the information about articles as dataframe.

```{r}
news_json$articles
```

### Question 3: Creating a Tidy Style Function for Querying NewsAPI .

#### Creating the function

Creating a function to query the API using a keyword, from_date and key as an input.

```{r}
news_api <- function(keyword, from_date, api_key) {
  base_url <- "https://newsapi.org/v2/top-headlines"
  
  # Construct the URL with parameters
  url <- modify_url(
    base_url,
    query = list(
      q = keyword,
      from = from_date,
      to = Sys.Date(),  # Today
      apiKey = api_key
    )
  )
  
  # Make the API request
  news <- GET(url)
  news_data <- rawToChar(news$content)
  news_json <- fromJSON(news_data)
  news_json$articles
  
}

```

#### Testing the function

Testing the function using the `keyword` = "Euro", `from_date` = "2024-05-25"

```{r}

news_api(keyword = "Euro",from_date = "2024-05-25",api_key = "e715bbaccd5849a0baf5c9fa6eb42a68") #Key Stored Locally
```

Trying with different keyword and different data

```{r}
news_api(keyword = "UEFA",from_date = "2024-06-05",api_key = "e715bbaccd5849a0baf5c9fa6eb42a68")
```
